Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-06-13 analysis_parameters.tex
cleaning and making progress with params

\section{Analysis Parameters}

% \subsection{Attenuation Relationship}

% The choice of the ground motion attenuation relationship, as a highly influential factor, is of a great importance in seismic hazard analysis. \citet{Zafarani2014} used 163 free-field acceleration time histories recorded at epicentral distance of up to 200 km from 32 earthquakes to investigate the predictive capabilities of the local, regional, and next generation attenuation (NGA) ground-motion prediction equations and determined their applicability for northern Iran. After evaluating different Ground Motion Prediction Equations, \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008} represented suitable performance for  PGA  with LLH (Log-Likelihood method) score of 1.54, 1.55, and 1.59. Mean of the mentioned attenuation relationships (not shown here), are very close together, especially at higher magnitude. Using \citet{Scherbaum2009} approach and \citet{Zafarani2014} coefficients we calculated the logic tree weights as 0.3376, 0.3354, and 0.3270, respectively. The epistemic uncertainty in attenuation relationship is being considered through using the logic tree approach. Having close mean value and similar logic tree weights, logic tree approach will have a minor effect in considering the epistemic uncertainty. In this study instead of using  different GMPEs we use the best model of attenuation relationship for PGA \citep{Zafarani2014} as well $\pm$ standard deviation. \\

% \noindent
% The attenuation relationship is:

% \begin{equation}
% \begin{split}
% ln\ (Y) = b_1 + b_2(M_w-6) + b_3( M_w-6)^{2}+  \\ 
% b_5ln\ r + b_V \ ln(V_S/V_A) \  with \  r= \sqrt{R{r^2_{cl} + h^2}}  
% \end{split}
% \end{equation}

@@

For each of the three seismic regions, Gutenberg-Richter parameters and maximum magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in each grid cells.\\
\noindent
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of \citet{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.

% \begin{figure*} [!ht]
% \centering
% \includegraphics[scale=0.3]{figures/pdf/figure-04} 
% \caption{The annual probability of exceeding the specified magnitude in different seismic tectonic region (a) Alborz, b) Azerbaijan, c) Central, d) Kopeh Dagh, e) Uniform f) Zagros). The margin display the positive and negative standard deviation.}
% \label{fig:annual_m}
% \end{figure*}

% \noindent

% \begin{table*}[!ht]
% \centering
% \caption{Seismicity parameters for three tectonic seismic regions.}
%     \begin{tabular}{ccccc}
%     ~                          &         $b-value$     &   $M_{max} $  Calculated & $M_{max}$ obseved \\ \hline
%     Azerbaijan           & 1.1   $\pm$ 0.03   &   7.93  $\pm$  0.34           &   7.7      			    \\ \hline
%     Alborz                  &1.03  $\pm$ 0.03   &   7.85  $\pm$  0.66           &   7.8          	            \\ \hline
%     Kopeh Dagh        & 0.89 $\pm$ 0.04   &   7.78  $\pm$  0.31           &   7.6          	            \\ \hline
%     Central Iran         & 0.95 $\pm$ 0.04   &   7.84  $\pm$  0.34           &   7.6                           \\ \hline
%     Zagros                 & 0.99 $\pm$ 0.02  &   7.47  $\pm$  0.26            &  7.4                            \\ \hline
%     Uniform Model    & 0.9    $\pm$ 0.02  &   7.87  $\pm$  0.26            &  7.8                            \\ 
%     \end{tabular}
%  \label{tab:b_value} 
% \end{table*}

@@

\begin{table*}[!ht]
\centering
\caption{Completeness threshold of tectonic seismic regions.}
\begin{tabular}{ccccccccc}
 ~           & Azerbaijan & Alborz & Kopeh Dagh & Central Iran & Zagros & North Iran  \\ \hline
3-4         & 1961          & 2005   & 2005             & 2005            & 2005    & 2005          \\ \hline
4-5         & 1961          & 1961   & 1950             & 1961            & 1961    & 1961          \\ \hline
5-6         & 1900          & 1900   & 1900             & 1854            & 1900    & 1900           \\ \hline
6-7         & 1581          & 1664   & 1673             & 1837            & 1853    & 1778           \\ \hline
7 $< $    & 1042          & -401    & 9                   & 762              & 1439    & -401            \\ 
\end{tabular}
\label{tab:completeness}
\end{table*}

>>>>

\section{Old Analysis Parameters Section}

@@

\textcolor{red}{
[Legacy text to be discussed with Naeem...]\\
}

\textcolor{blue}{
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of \citet{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.
}

@@

% \begin{table*}[!ht]
% \centering
% \caption{Completeness threshold of tectonic seismic regions.}
% \begin{tabular}{ccccccccc}
%  ~           & Azerbaijan & Alborz & Kopeh Dagh & Central Iran & Zagros & North Iran  \\ \hline
% 3-4         & 1961          & 2005   & 2005             & 2005            & 2005    & 2005          \\ \hline
% 4-5         & 1961          & 1961   & 1950             & 1961            & 1961    & 1961          \\ \hline
% 5-6         & 1900          & 1900   & 1900             & 1854            & 1900    & 1900           \\ \hline
% 6-7         & 1581          & 1664   & 1673             & 1837            & 1853    & 1778           \\ \hline
% 7 $< $    & 1042          & -401    & 9                   & 762              & 1439    & -401            \\ 
% \end{tabular}
% \label{tab:completeness}
% \end{table*}

### 2016-06-13 completeness.tex
cleaning and making progress with params

@@

\textcolor{red}{We will bring the completeness section here.}

>>>>

\begin{table*}[t]
    \centering
    \caption{Year thresholds and magnitude completeness for the seismic zones within the region of interest in this study.}
    \begin{tabular}{lccccrc}
    	\cline{2-7}                                              \\[-1.6ex]
                        & 3--4 & 4--5 & 5--6 & 6--7 & 
                                  \multicolumn{1}{c}{$>7$} &$M_c$\\[0.6ex]
        \hline                                                   \\[-1.6ex]
        Azerbaijan      & 1961 & 1961 & 1900 & 1581 & 1042 & 4.5 \\
        Alborz          & 2005 & 1950 & 1900 & 1664 &--401 & 4.4 \\
        Kopeh Dagh      & 2005 & 1950 & 1900 & 1673 &    9 & 4.5 \\
        Zagros          & 2005 & 1961 & 1900 & 1853 & 1439 & 4.4 \\
        Central-East    & 2005 & 1961 & 1854 & 1837 &  762 & 4.5 \\
        Uniform Model   & 2005 & 1961 & 1900 & 1778 &--401 & 4.5 \\[0.5ex]
        \hline 
    \end{tabular}
    \label{tab:params} 
\end{table*}

@@

An important property of a seismic catalog is its completeness, which is measured both in terms of the magnitude above w....

This parameter defines the minimum magnitude at which the seismic catalog of a region can be considered complete within a certain time period, and is used as input to the seismic analysis.

...which varies according to magnitude, is used as input during the seismic hazard analysis process. It basically consists in determining the point at which catalog data show a steady occurrence of events. There are multiple approaches to determine the completeness of a catalog. A simple approach used by \citet{Frankel1995} and others is to identify the year at which the rate of events becomes constant. In other words, if looking at a plot of cumulative number of events of a given magnitude range with respect to time, the catalog is assumed to be complete from the year at which moving forward the plot follows a straight line. A similar approach based on the maximum value of the first derivative of the frequency-magnitude curve is implemented in the ZMAP software \citep{Wiemer2001}.

One of the difficulties in determining completeness is the availability of data for a long-enough period of time for all magnitude ranges, especially for large magnitude earthquakes of low recurrence. \citet{Zare} provide the years for catalog completeness in their the original catalog for the different regions in their study on the seismicity of the Middle East. They used the

4.5, and 4.4 for 
Central Iran, and Zagros tectonic seismic regions, respectively. For the uniform model we used 4.5 as a completeness magnitude.

% 		& Azerbaijan  & Alborz & Kopeh Dagh & Central Iran & Zagros & North Iran  \\ \hline
% 3-4         & 1961  & 2005   & 2005       & 2005         & 2005   & 2005          \\ \hline
% 4-5         & 1961  & 1961   & 1950       & 1961         & 1961   & 1961          \\ \hline
% 5-6         & 1900  & 1900   & 1900       & 1854         & 1900   & 1900           \\ \hline
% 6-7         & 1581  & 1664   & 1673       & 1837         & 1853   & 1778           \\ \hline
% 7 $< $      & 1042  & -401   & 9          & 762          & 1439   & -401            \\

% For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor.  \citet{Frankel1995} plotted the cumulative number of events against time for different regions. He assumes that from the point that the line become linear, the catalog is complete. This approach is similar to the MAXC method of ZMAP software \citep{Wiemer2001}, where the completeness treshold is the maximum value of the first derivative of the frequency-magnitude curve.

% Using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and also frequency magnitude distribution approach in software ZMAP, \citet{Zare2014} reported the catalog completeness for the study regions.

% World widely large earthquakes were routinely located after increasing number of seismic stations establishment in the early 1900s \citep{Shearer2009}. Up to 1961 these data formed the early instrumental period. In 1961 the Worldwide Standardized Seismograph Network (WWSSN) was stablished. The record of these seismographs considerably improved the seismic catalogs in different part of the world \citep{Shearer2009}. Fig.~\ref{fig:completness_scatter} shows the magnitude distribution of events with respect to the time of occurrence of the events.

According to \citet{Frankel1995}, we made plots of the cumulative number of events against time for different regions' catalog.

% We pick the Mw 0.5 increments in magnitude to be able to compare the results with \citet{Zare2014}. In order to be able to compare the results with \citet{Zare2014} we merge the data of Azerbaijan and Alborz seismic regions.  Fig.~\ref{fig:completness_compare_zare_2014_Az_Al} shows the completeness of catalog for earthquakes with different magnitude range in Azerbaijan-Alborz region. According to this figure, midrange magnitudes ($ 4 < Mw < 6 $) fairly obey the network developments in 1900 and 1961. The completeness thresholds for each magnitude range which are reported by \citet{Zare2014} are shown by dashed green line.

% In this study we follow the \citet{Frankel1995} approach to determine the completeness of each region. Even though the approach used by \citet{Frankel1995} will lead to the conservative results, we make sure that we don't use a period of time without knowing the complete number of events. Determining the completeness of the catalog is very sensitive to the data. Converting earthquake magnitudes from different scales to moment magnitude obviously has some error. Having broader range of magnitude will help to minimize these sort of error. In this study we consider the magnitude intervals for completeness study as $Mw = 1$.

### 2016-06-13 composition.tex
cleaning and making progress with params

A fundamental step in regional seismic hazard analysis is the composition of a reliable seismic catalog that includes both historically documented and instrumentally recorded earthquakes. Nowadays, instrumental datasets are easily accessible through international, regional, and local agencies. Historical data, on the other hand, lends itself to subjective interpretation. Notable previous efforts to compose uniform earthquake catalogs for Iran include but are not limited to \citet{Ambraseys_1982_Book}, \citet{moinfar1994}, and \citet{Berberian_1995_Tech}. These studies, however, differ from each other in their use of magnitude scales or magnitude conversion rules, and the extent to which they accept historical accounts as valid data.

>>>>

The first fundamental step in regional seismic hazard analysis is the composition of a reliable seismic catalog that includes both historically documented and instrumentally recorded earthquakes. Nowadays, instrumental datasets are easily accessible through international, regional, and local agencies. Historical data, on the other hand, lends itself to subjective interpretation. Notable previous efforts to compose uniform earthquake catalogs for Iran include but are not limited to \citet{Ambraseys_1982_Book}, \citet{moinfar1994}, and \citet{Berberian_1995_Tech}. These studies, however, differ from each other in their use of magnitude scales or magnitude conversion rules, and the extent to which they accept historical accounts as valid data.

### 2016-06-13 region.tex
cleaning and making progress with params

\begin{figure*}[t] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-02}
	\caption{(a) Main fault line and fault systems, and (b) notable earthquakes in the region of interest. White lines delimit the three major seismic zones in northern Iran. Epicentral locations are indicated with stars. The labels next to each star or linked with a line to avoid congestion show the magnitude and year of the earthquake.}
	\label{fig:selected}
\end{figure*}

@@

Transitioning from the Azerbaijan to the Alborz seismic zone, the next fault system runs along the Rocks of the Talesh mountains, to the northwest of the Alborz Mountains. Faults in this area have also ruptured recently \citep{Berberian1999}. The $M_w$ 7.4 1990 Manjil-Rudbar earthquake, for instance, caused numerous deaths and significant damage to the region in the south Caspian depression. At the central area of the Alborz seismic zone, near Tehran, the region is characterized by active reverse faults in the North Tehran Thrust (NTT) system (see Fig.~\ref{fig:selected}a for reference). These faults are parallel to the northwest-trending structural gain of the Alborz Mountains belt, and include the south-dipping partially blind faults of Davudieh, Shian, and Bagh-E Feyz. According to \citet{Ambraseys2005}, Tehran---or the ancient city of Rey---has been devastated in multiple occasions due to severe earthquakes in this area. Historical accounts document the occurrence of multiple $M>7$ earthquakes in 743, 958, 1177, 1665, and 1830. The 1830 earthquake, in particular, ruptured the central segments of the NTT, adjacent to the Mosha fault.

>>>>

\begin{figure*}[t] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-02}
	\caption{(a) Main fault line and fault systems, and (b) notable earthquakes in the region of interest. White lines delimit the three major seismic zones in northern Iran. Epicentral locations are indicated with stars. The labels next to each star or linked with a line to avoid congestion show the magnitude and year of the earthquake. The borderlines of the seismic zones shown in Fig.~\ref{fig:iran} are shown in the background in white, along with the light gray shaded relief.}
	\label{fig:selected}
\end{figure*}

@@

Transitioning from the Azerbaijan to the Alborz seismic zone, the next fault system runs along the Rocks of the Talesh mountains, to the northwest of the Alborz Mountains. Faults in this area have also ruptured recently \citep{Berberian1999}. The $M_w$ 7.4 1990 Manjil-Rudbar earthquake, for instance, caused numerous deaths and significant damage to the region in the south Caspian depression. At the central area of the Alborz seismic zone, near Tehran, the region is characterized by active reverse faults in the North Tehran Thrust (NTT) system (see Fig.~\ref{fig:selected}a for reference). These faults are parallel to the northwest-trending structural gain of the Alborz Mountains belt, and include the south-dipping partially blind faults of Davudieh, Shian, and Bagh-E Feyz. According to \citet{Ambraseys_1982_Book}, Tehran---or the ancient city of Rey---has been devastated in multiple occasions due to severe earthquakes in this area. Historical accounts document the occurrence of multiple $M>7$ earthquakes in 743, 958, 1177, 1665, and 1830. The 1830 earthquake, in particular, ruptured the central segments of the NTT, adjacent to the Mosha fault.
